Hydrodynamics of a Semipermeable Vesicle Under Flow and Confinement

### 2020-12-30 vesicle_code/docs/permeableVes/main.tex
working on supp material

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semipermeable Vesicle in a Contracting Geometry} 
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} A contracting geometry
  similar to the experimental device in~\cite{wu2015critical}. The
  channel width varies from $18\mu$m to $1.8\mu$m, and it repeats every
  $45\mu$m. The fluid flux along the red regions is into the vesicle,
  and the fluid flux along the blue regions is out of the vesicle.
  (a--c): The ability for a semipermeable vesicle to pass through the
  contraction depends on the semipermeability rate. (d--f): The ability
  for a semipermeable vesicle to pass through the contraction depends
  on the flow rate. The colored vesicles show the tension distribution
  in N/m of the circled vesicles. Throughout the entire simulation,
  these vesicle configurations have the largest tension.}
\end{figure}
Squeezing a cell or an elastic vesicle through a slit in microfluidics
has gained popularity for its application in stress-induced release of
macromolecules~\cite{ShareiEtAl2013_PNAS, Pak2015_PNAS,
ZhangShenHoganBarakatMisbah2018_BJ, LuoBai2019_PoF}. Red blood cells go
through submicron slits many times during their lifetime, and such
process selects healthy, flexible red blood cells to
survive~\cite{wu2015critical,LuPeng2019_PoF}. Here we investigate the
role of membrane permeability to water when a semipermeable vesicle is
squeezed through a contracting microfluidic channel similar
to~\cite{wu2015critical}, where the width of the channel gradually
decreases from $18 \mu$m to $1.8 \mu$m. The channel then immediately
opens again to $18 \mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant
maximum flow rate of $U_{\max}$. We initialize a semipermeable vesicle
with $\alpha = 0.65$ that spans $10 \mu$m of the total channel width and
investigate the effects of the membrane permeability to water and the
maximum flow velocity on whether a vesicle can successfully pass through
the slit. We color the vesicle according to the sign of its flux---red
for influx and blue for efflux.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Numerical Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using a boundary integral equation formulation, the unknowns are the
vesicle's position and tension and the unknown density function for
confined flows. Both $\gamma$ and $\partial\Omega$ are discretized at
$N$ and $N_\text{wall}$ points. Derivatives are computed with Fourier
differentiation, and the eighth-order quadrature~\cite{alp1999} is
applied to the weakly singular single-layer potential. For confined
geometries, the trapezoid rule is used to approximate the double-layer
potential since its kernel is smooth and periodic. Nearly-singular
integrals are computed with an interpolation-based quadrature
rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semiimplicitly~\cite{vee-gue-zor-bir2009}. We introduce the notation
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
for the differential and projection operators. Then, the
governing equations~\eqref{eqn:vesVelocity} are
\begin{align}
  \dot{\xx} &= \uu_\infty(\xx) + \beta P(B\xx + T\sigma)
  + \SS (B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
We note that the integral and differential operators, which all depend
on the vesicle shape, are non-linear with respect to the vesicle
shape. We construct a time stepping scheme by letting $B^N$ be the
bending operator due to the vesicle configuration at time $t^N$, and use
similar notation for the other operators. Then, a first-order
semiimplicit time stepping method of~\eqref{eqn:vesVelocity} is
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\triangle t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right),  \\
  D^N\xx^{N+1} &= 1,
\end{alignat}
and this is a linear system that we solve with matrix-free GMRES.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Vesicle Area}
In the absence of an imposed flow, the steady state vesicle shape and a
differential equation for the area of the vesicle can be derived. We
start by following the analysis of Veerapaneni {\em et
al.}~\cite{vee-raj-bir-pur2009} by computing the steady state shape of a
semipermeable vesicle in a quiescent flow. The steady state shape is a
stationary point of the Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note the absence
of the pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curveODE}
\end{align}
Assuming $\kappa_s \neq 0$,~\eqref{eqn:curveODE} is separable, and can
be solved analytically in terms of the {\em EllipticF} function using
Mathematica. Unlike in the impermeable case, the solution
of~\eqref{eqn:curveODE} is not periodic unless $\kappa(s) = \kappa_0$.
Therefore, the steady state shape of a semipermeable vesicle in a
quiescent flow is a circle with radius $L/2\pi$ and constant tension
$\Lambda = 2\pi^2/L^2.$ Note that in contrast to a impermeable vesicle,
the tension is positive rather than negative.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semipermeable Vesicle in a Contracting Geometry} 
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \else
  \includegraphics{figures/contractingComposite.pdf}
  \fi
  \caption{\label{fig:contractingComposite} A contracting geometry
  similar to the experimental device in~\cite{wu2015critical}. The
  channel width varies from $18\mu$m to $1.8\mu$m, and it repeats every
  $45\mu$m. The fluid flux along the red regions is into the vesicle,
  and the fluid flux along the blue regions is out of the vesicle.
  (a--c): The ability for a semipermeable vesicle to pass through the
  contraction depends on the semipermeability rate. (d--f): The ability
  for a semipermeable vesicle to pass through the contraction depends
  on the flow rate. The colored vesicles show the tension distribution
  in N/m of the circled vesicles. Throughout the entire simulation,
  these vesicle configurations have the largest tension.}
\end{figure}

Squeezing a cell or an elastic vesicle through a slit in microfluidics
has gained popularity for its application in stress-induced release of
macromolecules~\cite{ShareiEtAl2013_PNAS, Pak2015_PNAS,
ZhangShenHoganBarakatMisbah2018_BJ, LuoBai2019_PoF}. Red blood cells go
through submicron slits many times during their lifetime, and such
process selects healthy, flexible red blood cells to
survive~\cite{wu2015critical,LuPeng2019_PoF}. Here we investigate the
role of membrane permeability to water when a semipermeable vesicle is
squeezed through a contracting microfluidic channel similar
to~\cite{wu2015critical}, where the width of the channel gradually
decreases from $18 \mu$m to $1.8 \mu$m. The channel then immediately
opens again to $18 \mu$m (Figure~\ref{fig:contractingComposite}). A
Poiseuille flow is imposed at the inlet and outlet with a constant
maximum flow rate of $U_{\max}$. We initialize a semipermeable vesicle
with $\alpha = 0.65$ that spans $10 \mu$m of the total channel width and
investigate the effects of the membrane permeability to water and the
maximum flow velocity on whether a vesicle can successfully pass through
the slit. We color the vesicle according to the sign of its flux---red
for influx and blue for efflux.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Numerical Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Space}
By using a boundary integral equation formulation, the unknowns are the
vesicle's position and tension and the unknown density function for
confined flows. Both $\gamma$ and $\partial\Omega$ are discretized at
$N$ and $N_\text{wall}$ points, respectively. Derivatives are computed
with Fourier differentiation, and the eighth-order
quadrature~\cite{alp1999} is applied to the weakly singular single-layer
potential. For confined geometries, the trapezoid rule is used to
approximate the double-layer potential since its kernel is smooth and
periodic. Nearly-singular integrals are computed with an
interpolation-based quadrature rule~\cite{qua-bir2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Discretization in Time}
To eliminate stringent time step restrictions, we modify a time stepping
method that discretizes high-order derivatives
semi-implicitly~\cite{vee-gue-zor-bir2009}. We introduce the notation
\begin{alignat}{3}
  &B[\xx]\ff = -\frac{d^4}{ds^4} \ff,  \qquad
  &&T[\xx]\sigma = (\sigma \xx_s)_s, \\
  &D[\xx]\ff = \xx_s \cdot \ff_s, 
  &&P[\xx]\ff = (\ff \cdot \nn) \nn,
\end{alignat}
for the differential and projection operators. Then, the
governing equations~\eqref{eqn:vesVelocity} are
\begin{align}
  \dot{\xx} &= \uu_\infty(\xx) + \beta P(B\xx + T\sigma)
  + \SS (B\xx + T\sigma), \quad
  D \dot{\xx} = 0.
\end{align}
We note that the integral and differential operators, which all depend
on the vesicle shape, are non-linear with respect to the vesicle
shape. We construct a time stepping scheme by letting $B^N$ be the
bending operator due to the vesicle configuration at time $t^N$, and use
similar notation for the other operators. Then, a first-order
semi-implicit time stepping method of~\eqref{eqn:vesVelocity} is
\begin{alignat}{3}  
  \frac{\xx^{N+1} - \xx^N}{\triangle t} &= \uu_\infty(\xx^N) 
  &&+ \beta P^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right) \nonumber \\
  & &&+ \SS^N\left(B^N\xx^{N+1} + T^N\sigma^{N+1}\right),  \\
  D^N\xx^{N+1} &= 1,
\end{alignat}
and this linear system is solved with matrix-free GMRES.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supplementary Material: Vesicle Area}
In the absence of an imposed flow, the steady state vesicle shape and a
differential equation for the area of the vesicle can be derived. We
start by following the analysis of Veerapaneni {\em et
al.}~\cite{vee-raj-bir-pur2009} by computing the steady state shape of a
semipermeable vesicle in a quiescent flow. The steady state shape is a
stationary point of the Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where the Lagrange multiplier $\Lambda$ is the tension. Note the absence
of the pressure Lagrange multiplier since the vesicle's area fluctuates.
Taking the variation of $\mathcal{L}$ with respect to $\gamma$ and
setting it to zero, the steady state shape satisfies $\kappa_{ss} +
\frac{1}{2}\kappa^3 - \Lambda \kappa = 0$, with initial conditions
$\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.  Integrating once, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \frac{\kappa_0^4}{8} - 
    \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curveODE}
\end{align}
Assuming $\kappa_s \neq 0$,~\eqref{eqn:curveODE} is separable, and can
be solved analytically in terms of the {\em EllipticF} function using
Mathematica. Unlike in the impermeable case, the solution
of~\eqref{eqn:curveODE} is not periodic unless $\kappa(s) = \kappa_0$.
Therefore, the steady state shape of a semipermeable vesicle in a
quiescent flow is a circle with radius $L/2\pi$ and constant tension
$\Lambda = \kappa^2/2 = 2\pi^2/L^2.$ Note that in contrast to a
impermeable vesicle, the tension is positive rather than negative.
